\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-25 paper.qmd
Updated PBMC3k layouts lables to have the consistency

@@

nldr2 <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour=clr_choice) +
  interior_annotation("b")

nldr2c <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#f781bf') +
  interior_annotation("b")

#umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")
umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
nldr3 <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour=clr_choice) +
  interior_annotation("c")

nldr3c <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#999999') +
  interior_annotation("c")

@@

```{r}
#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h). Layout f is universally poor. Layouts a, b, g, h that show large separations between clusters are universally suboptimal. Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases. The choice of best is between layouts c and e, that have small separations between oddly shaped clusters. Layout e is the best choice."
#| label: fig-pbmc-mse
#| fig-pos: H
#| out-height: 100%

@@

```{r}
#| echo: false
#| fig-cap: "Model in \\gD{}, on the layout a and c, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$). Layout a shows three-well separated clusters with big separation, while layout c shows close clusters. In $9\\text{-}D$, the data shows close clusters. This is the reason where b1, b2, and b3 shows connected edges between the clusters, and also proved by @fig-pbmc-mse as the best layout, which used the same NLDR method. By viewing the fitted model in $9\\text{-}D$, help to see some unobserved pattern of the data (i) densed points, and (ii) non-linear clusters."
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 15
#| out-height: 80%

>>>>

umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_12_min_dist_0.99.rds")
nldr2 <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour=clr_choice) +
  interior_annotation("b")

nldr2c <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#999999') +
  interior_annotation("b")

@@

nldr3 <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour=clr_choice) +
  interior_annotation("c")

nldr3c <- umap_pbmc |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#f781bf') +
  interior_annotation("c")

#umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")

@@

```{r}
#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h). Layout f is universally poor. Layouts a, c, g, h that show large separations between clusters are universally suboptimal. Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases. The choice of best is between layouts b and e, that have small separations between oddly shaped clusters. Layout e is the best choice."
#| label: fig-pbmc-mse
#| fig-pos: H
#| out-height: 100%

@@

```{r}
#| echo: false
#| fig-cap: "Model in \\gD{}, on the layout a and b, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$). Layout a shows three-well separated clusters with big separation, while layout b shows close clusters. In $9\\text{-}D$, the data shows close clusters. This is the reason where b1, b2, and b3 shows connected edges between the clusters, and also proved by @fig-pbmc-mse as the best layout, which used the same NLDR method. By viewing the fitted model in $9\\text{-}D$, help to see some unobserved pattern of the data (i) densed points, and (ii) non-linear clusters."
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 15
#| out-height: 80%

### 2024-09-25 paper.tex
Updated PBMC3k layouts lables to have the consistency

\caption{\label{fig-pbmc-mse}Assessing which of the 8 NLDR layouts on
the PBMC3k data (shown in Figure~\ref{fig-NLDR-variety}) is the better
representation using MSE for varying binwidth (\(a_1\)). Colour used for
the lines and points in the left plot and in the scatterplots represents
NLDR layout (a-h). Layout f is universally poor. Layouts a, b, g, h that
show large separations between clusters are universally suboptimal.
Layout d with little separation performs well at tiny binwidth (where
most points are in their own bin) and poorly as binwidth increases. The
choice of best is between layouts c and e, that have small separations
between oddly shaped clusters. Layout e is the best choice.}

@@

\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the layout
a and c, and two views of the fit in projections from \(9\text{-}D\),
for the PBMC3k data (\(n =  2622\) and \(p = 9\)). Layout a shows
three-well separated clusters with big separation, while layout c shows
close clusters. In \(9\text{-}D\), the data shows close clusters. This
is the reason where b1, b2, and b3 shows connected edges between the
clusters, and also proved by Figure~\ref{fig-pbmc-mse} as the best
layout, which used the same NLDR method. By viewing the fitted model in
\(9\text{-}D\), help to see some unobserved pattern of the data (i)
densed points, and (ii) non-linear clusters.}

>>>>

\caption{\label{fig-pbmc-mse}Assessing which of the 8 NLDR layouts on
the PBMC3k data (shown in Figure~\ref{fig-NLDR-variety}) is the better
representation using MSE for varying binwidth (\(a_1\)). Colour used for
the lines and points in the left plot and in the scatterplots represents
NLDR layout (a-h). Layout f is universally poor. Layouts a, c, g, h that
show large separations between clusters are universally suboptimal.
Layout d with little separation performs well at tiny binwidth (where
most points are in their own bin) and poorly as binwidth increases. The
choice of best is between layouts b and e, that have small separations
between oddly shaped clusters. Layout e is the best choice.}

@@

\caption{\label{fig-model-pbmc-author-proj}Model in \gD{}, on the layout
a and b, and two views of the fit in projections from \(9\text{-}D\),
for the PBMC3k data (\(n =  2622\) and \(p = 9\)). Layout a shows
three-well separated clusters with big separation, while layout b shows
close clusters. In \(9\text{-}D\), the data shows close clusters. This
is the reason where b1, b2, and b3 shows connected edges between the
clusters, and also proved by Figure~\ref{fig-pbmc-mse} as the best
layout, which used the same NLDR method. By viewing the fitted model in
\(9\text{-}D\), help to see some unobserved pattern of the data (i)
densed points, and (ii) non-linear clusters.}
